\documentclass[pdf]{beamer}
\mode<presentation>{}

\usepackage{tipa}
\usepackage{kpfonts}
\usepackage{xcolor}
 \definecolor{dgreen}{rgb}{0.,0.6,0.}
\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}

\newcommand{\slidd}[5] {
    \begin{frame}{}
            \begin{center}
                \vfill
                \textbf{#1}\\\vspace{1em}
                {\small #4}\\\vspace{1em}
                \textipa{#3}\\\vspace{1em}
                {\scriptsize background absolute tense: #5}\\\vfill
                \emph{\Large #2}\\\vfill
            \end{center}
        \end{frame} 
        }


        \newcommand{\Sk}[1]{.\textsc{#1}}
        \newcommand{\sK}[1]{\textsc{#1}.}

\begin{document}





\slidd{{N Modar forba orodan ekagardaekat kyttkattk,}}{{I met Mydar around the day 54}}{{ \s{n}"mOdar "fOrba oro"dan Ekagard\t*{ae}"kat @k1jt:"kat:k}}{ I\Sk{1s}\Sk{erg} \sK{abs}Mydar approximately \sK{loc}day 54\Sk{gen} meet\Sk{ind}\Sk{ant} }{present}


\slidd{kym rda ngon oKorma rarodany pyka.}{when I stationed in Korma for a few days.}{k\s{m}"rdaNon o"kOrma raro"da\textltailn 1 "pj1ka}{while (set time to past) I\Sk{1s}\Sk{erg} \sK{loc}Korma \sK{abl}day\Sk{pl} few}{past}


\slidd{Moshl net nar fen ong kattk}{She was brought to me by her mother,}{"mOS\s{l}nEt nar"fEn ON"kat:k }{mother her\Sk{3s}\Sk{gen} \textsc{fem} \sK{loc}I\Sk{1s} bring\Sk{ind}\Sk{ant}}{past}


\slidd{dyp shledd fettfok nadd formadat omoshl.}{after she had noticed the daughter was acting strangely.}{d1jp SlEd: "fEt:fok nad:@ forma"dat o"mOS\s{l}}{ after daughter\Sk{erg} strangely (seems like) behave\Sk{ind}\Sk{pr} mother\Sk{dat}  }{past}


\slidd{Fattfyt n Modar otaryng nget kat.}{I immedately brought Mydar to my tent.}{"fat:fj1t \s{n}"mOdar ota"rjWN @Net kat}{immediately I\Sk{1s}{erg} \sK{abs}Mydar \sK{loc}tent my\Sk{1s}\Sk{gen} bring\Sk{ind}\Sk{pr}}{past}


\slidd{Gegord dhana, ngon fattfyt bag goma net rdonaroda eat, baberbef yrk gagyrbef ena.}{Upon examination, I noticed immediately her skin was flaky, in the arms and legs.}{"gEgord "D:ana Non fat:fj1t bag "gOma nEt rdona"rOda \t*{eat}, ba"bErbef \s{r}k @ga"g\s{r}bef "Ena } {\sK{abl}examination after, I immediately \sK{abs}[skin her\Sk{3s}\Sk{gen} flaky (was)] see\Sk{ind}\Sk{pr}, \sK{abl}arm\Sk{dual} and \sK{abl}leg\Sk{dual} on}{past}

\slidd{Nor fa bag oyrabad esh nar nedhattk damatam}{She confessed to having swum in the lake.}{ "nOrfa bag o.1ra"bad ES"nar ne"D:at:k dama"tam}{ she\Sk{3s}\Sk{erg} \textsc{fem} \sK{abs}[\sK{loc}lake in she\Sk{abs} swim\Sk{ind}\Sk{ant} ] confess\Sk{ind}\Sk{pr} }{past}


\slidd{Nor bag ybabake onorg desheshot sasao gym pyrdek bordam}{She said she wanted to silence the spirits that would whisper to her through her portal.}{"nOrbag iba"bake "Onorg de"SESot sa"s\t*{ao} g\s{m}"p\s{r}dek "bOrdam }{she\Sk{erg} \sK{abs}[\sK{plural}\sK{abs}spirit her\Sk{dat} \sK{patpart}whisper\Sk{subj}\Sk{pr} \sK{abl}silence into bring\Sk{deo}{ind}] tell\Sk{ind}\Sk{pr} }{past}


\slidd{Otta dhlashlaryb, kyrk maegadd shlyddat kedagyardat}{Her nose was dry, so I diagnosed a necrotic imbalance.}{Ot:a Dla"Slar1jb k\s{r}k m\t*{ae}"gad: Sl1"d:at kedagja"rdat}{nose\Sk{erg} dry (is), therefore (I) \sK{abs}imbalance blue\Sk{gen} declare\Sk{ind}\Sk{pr}}{past}



\slidd{Nar ttyttakyba kat ong bo kyta-rdyad.}{She never came back for the second visit.}{nar @t:1t:a"k1jba ka"tON bO k1ta"rd\t*{Wa}d}{she\Sk{erg} \sK{abl}visit two\Sk{gen} (for) \sK{loc}me to \sK{not}go\Sk{ind}\Sk{post}  }{past}




\end{document}
